
---
title: "Practical Machine Learning - Prediction Assignment"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---


# Executive Summary

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity. The main objective is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to quantify quantify how well they do  a particular activity

# Exploratory Analysis

```{r}
library(caret);library(rattle); library(rpart.plot);library(doParallel)
setwd("C:\\Raphael\\estudos\\datasciencecoursera\\Machine Learning")
# download.file('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv', 'pml-training.csv')
# download.file('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv','pml-testing.csv' )

train_file = read.csv('pml-training.csv', na.strings=c("NA","#DIV/0!", ""))
test_file = read.csv('pml-testing.csv' , na.strings=c("NA", "#DIV/0!", ""))

str(train_file)
table(train_file$classe)

```

# Preparing data

In order to clean data, I've decided to exclude all variables with NA's and other with no direct relationship to "classe".

```{r}

selected = which((colSums(!is.na(train_file)) >= nrow(train_file)))
train_data = train_file[,selected]
test_data = test_file[,selected]

test_data = test_data[-ncol(test_data)]
test_data$new_window = factor(test_data$new_window, levels=c("no","yes"))

train_data = train_data[,-c(1,5)]
test_data = test_data[,-c(1,5)]

names(selected)

```


# Training data

70% train, 30% test

```{r}

set.seed(10)
inTrain = createDataPartition(train_data$classe, p = 0.7, list = FALSE)
training = train_data[inTrain, ]
testing = train_data[-inTrain, ]

```

# Random Forest

```{r}

class = training$classe
data = training[-ncol(training)]

registerDoParallel()
modFit <- train(data, class, method="parRF", tuneGrid=data.frame(mtry=3), trControl=trainControl(method="none"))

modFit
modFit$finalModel


test_predict <- predict(modFit, newdata=testing)
confMatrix <- confusionMatrix(test_predict,testing$classe)
confMatrix
confMatrix$overall[1]

```

The model seems to predict quite well, once the Accuracy is about 99.7%.


# Decision Tree

```{r}

modFit1 <- rpart(classe ~ ., data=training, method="class")
fancyRpartPlot(modFit1)

test_predict_dt <- predict(modFit1, testing, type = "class")
confusionMatrix(test_predict_dt, testing$classe)

confMatrix$overall[1]

```


# Save files to submit

```{r}

pml_write_files = function(x){
        n = length(x)
        for(i in 1:n){
                filename = paste0("problem_id_",i,".txt")
                write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
        }
}

answers <- predict(modFit, test_data)
pml_write_files(answers)

```